---
title: "Tooth Growth"
author: "jagdeep s sihota"
output: pdf_document
---

## ToothGrowth: Data Analysis

### 1. Load the ToothGrowth data : 
```{r}
library(datasets) 
library(ggplot2)
data(ToothGrowth) 
summary(ToothGrowth)

ggplot(ToothGrowth, aes(x=supp, y=len)) + geom_boxplot()
ggplot(ToothGrowth, aes(x=interaction(supp, dose), y=len)) + geom_boxplot()
```

#### Basic summary of the data

Data shows administering larger doses of vitamin C resulted in, on average, longer teeth, but we can't figure out which method of administering the vitamin is more effective.

### 3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose:

#### 3.1. Tooth Growth by Suppliment:

Average increased tooth length for the OJ group versus the VC group:
```{r}
OJTest <- ToothGrowth[ToothGrowth$supp=="OJ", "len"]
VCTest <- ToothGrowth[ToothGrowth$supp=="VC", "len"]
lengthChange=OJTest - VCTest 
``` 
The group treated with OJ has higher mean length of `r  mean(OJTest)` vs the VC group having the mean of about `r  mean(VCTest)`. Is this  mean administering OJ makes average tooth growth bigger?

Setting up the null and alternate hypotheses: <b>Ho:</b> lengthChange=0 VS. <b>Ha:</b> lengthChange>0

```{r} 
lower_bound_ci=mean(lengthChange) - 1.96*sd(lengthChange)/sqrt(length(lengthChange))
upper_bound_ci=mean(lengthChange) + 1.96*sd(lengthChange)/sqrt(length(lengthChange))

``` 
We reject the null hypothesis becasue the confidence interval (`r lower_bound_ci`,`r upper_bound_ci`) does not include 0 - the mean length change for null hypothesis



#### 3.2. Tooth Growth by Dose:

##### Tooth Growth by Suppliment when dosage is (0.5,1):
```{r}
lowOJ <- ToothGrowth[ToothGrowth$supp=="OJ" & ToothGrowth$dose %in% c(0.5,1),"len"] 
lowVC <- ToothGrowth[ToothGrowth$supp=="VC" & ToothGrowth$dose %in% c(0.5,1),"len"]
```
Confidence interval and test the hypothesis: 
```{r}
alpha<-0.05; 
th<-t.test(lowOJ,lowVC); 
```

For t>0, confidence interval limits >0, <b>Ho:</b>meanOfOJ = meanOfVC VS. <b>Ha</b>: meanOfOJ > meanOfVC

Ho is rejected, because TS(t = `r th$statistic`) > qt(1-alpha,`r th$parameter`)( = `r qt(1-alpha,th$parameter)`). 

##### Tooth Growth by Suppliment when dosage is 2:
```{r}
highOJ<- ToothGrowth[ToothGrowth$supp=="OJ" & ToothGrowth$dose==2,"len"] 
highVC<- ToothGrowth[ToothGrowth$supp=="VC" & ToothGrowth$dose==2,"len"]
th<-t.test(highOJ,highVC); 
```
t<0, test the hypothes: <b>Ho:</b>meanOfOJ = meanOfVC versus <b>Ha:</b> meanOfOJ < meanOfVC

Ho is rejected, because TS(t = `r th$statistic`) < qt(1-alpha,`r th$parameter`)( = `r qt(1-alpha/2,th$parameter)`). 
 
### 4. State your conclusions and the assumptions needed for your conclusions:

#### Assumptions:

* the hypothesis is: both supp have the same impact on tooth length.
* the alternative hypothesis is: OJ can promote the growth of toothlength than VC.
* significance level: 0.05

#### Conclusions:

* OJ have more impact on the growth of toothlength under dose equals to 0.5 or 1.But can't get this conclusion when dose equals to 2.